Alleviating Topology Mismatch in Distributed Overlay Networks: A Survey

### 2014-01-27 structured.tex
2012.12.23 (Mema) Text reworked.

\section{Decentralized Structured Architectures}
\label{section:structured}

This section focuses on the structured P2P architectures. The following
paragraphs contain a thorough discussion and analysis of the algorithms proposed
in the literature for dealing with the topology mismatch problem in such
architectures. In structured P2P systems, the construction of the routing tables
in every participating node determines the efficiency of message forwarding.
Therefore, routing tables that represent the underlying physical structure well
can achieve increased performance. For this reason, the proposed methods for
topology mismatch problem basically use different levels of proximity
information to optimize them towards this direction. The algorithms are here
categorized based on the previous work done by \cite{CDHR2002,CDCR2002,RSS2002}.

@@

\textit{Global Soft-State} \cite{XTZ2003} builds a global map to help
choose shorter routing paths, combining the landmark binning method and small
scale distance probes to reveal the proximity properties of the underlying
network to the overlay. %Figure \ref{fig:global_softstate} presents the landmark
%space and the mapping of the position of the nodes.
This global view of the
state is made available to all nodes in order to help them find the best way to
route their messages. \textit{Global Soft-State} operates in two main stages,
generation and use of the proximity information. For generating the proximity
information a hybrid approach is proposed, which uses landmark clustering as a
preprocessing step in order to select a number of potential nearest neighbor
candidates and then refine the selection by incorporating a
round-trip-time-based scheme to ultimately choose the closest node. For using
the proximity information, the algorithm chooses a different path from the
classic gossiping approaches for constructing and maintaining the overlay. It is
based on landmark clustering for strategic placement of proximity information on
the overlay enabling any node to access such information using a landmark number
that reflects its physical position in the network. For various logical
regions\footnote{This might be a high-order zone in the eCAN\cite{xu_ecan_2002}
context or a set of nodes sharing a particular prefix in overlays such as
Pastry\cite{antony_pastry_2001}.} maps of physical information are built and
published where each node may appear in a maximum of $log\left( N \right)$ such
maps. To dynamically adapt to changing network conditions, a node subscribes to
relevant \emph{soft states} that utilize a notification system in order to
initialize any necessary neighbor re-selection.

Maintaining several host states at different layers, makes any content migration
costly. Additionally, the method does not make any continuing effort to remap
the overlay structure after a node successfully joins, in order to adapt its
state to any condition change. Although this approach greatly reduces the
routing latency to far nodes, it is unable to dynamically identify nodes that
are close to routers and gateways in order to construct the secondary overlay.
Nevertheless, static recognition of such nodes is currently done based on BGP
reports and pre-chosen landmarks, sacrificing the self-organizing attribute of
traditional DHTs.

@@

The bootstrap phase of Mithos starts with a subset of existing members as the
first set of candidates, and while the iteratively closest nodes are detected by
probing the neighborhood, the candidate neighbor list is updated. In order to
avoid a local minima, Mithos probes all the neighbors within a two-hop distance
from the current minimum before concluding the process.

After finding its first neighbor, the new coming peer is assigned an ID using
information gathered during the iterative neighbor selection phase. Virtual
coordinates are assigned to the new coming peer by using the distances of two
closest nodes and their neighbors, so that Euclidean distances between the node
and all known hosts predict the network latency between them
\cite{cox_vivaldi_2004}. The benefit of these synthetic coordinates are that
they are explicitly used as the node's ID, and distance computations among nodes
can be done in ID space without requiring physical measurements.

The last step of the algorithm is the interconnection among neighbors. Mithos
uses a \emph{quadrant-based} mechanism according to which each node establishes
a link to the closest neighbor in each quadrant. During forwarding, the next
hop is performed towards a neighbor in the same quadrant as the final
destination. The new coming node may not know of other neighbors in all
quadrants, therefore, the node first identifies neighbors in all quadrants
using a mechanism based on ideas similar to a perimeter walk\footnote{Used in
Greedy Perimeter Stateless Routing (GPSR) protocol.} and then improves the
results using parallel path processing by taking into account further geometric
properties of node relationships.

One major limitation of \textit{Mithos} is that the protocol cannot handle
dynamic arrivals and departures from the network, as is happens in mobile
networks.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{LAPTOP}
\emph{LAPTOP} \cite{WLH2007}, organizes the overlay into a tree-based hierarchy
with main focus on reducing hops during message routing as well as minimizing
maintenance overhead. Additionally, a caching scheme is also incorporated so as
to further reduce routing table update costs. The authors theoretically show
that in LAPTOP routing path length is bounded by $O(log_d N)$ and node joining
and leaving in the overlay network is bounded by $O\left( d log_d N \right)$
hops in a balanced overlay tree, where $N$ is the number of nodes, and $d$ is
the maximum degree of each node. LAPTOP implements the geographical layout
approach  and constructs its layout in a self-organizing and efficient fashion.
The physical distances are estimated with round trip times to a few existing
nodes in the overlay network.

Each node is assigned a dot-formated address (e.g 1.3.4). Each octet ranges from
$1$ to $d$, where $d$ is the maximum degree of the nodes. The assignment process
is done by appending a unique octet to the address of each nodes parent, while
root node is assigned address $1$.

@@

The routing scheme is similar to the longest-prefix IP matching scheme. At each
forwarding hop, any message travels up the tree until the first common ancestor
of both source and destination nodes is reached and then starts descending to
arrive to its target. During tree traversals, special entries in the routing
tables, called \emph{routing cache}, are maintained in order to increase routing
efficiency and achieve finer load balance. Caching enables a node to forward a
message to a better longest-prefix match than that of its direct ancestor making
a larger, quicker and more cost effective step through the overlay and toward
the destination. To improve scalability, the number of children nodes and the
size of the routing cache are limited. In terms of overlay maintenance, LAPTOP
incorporates a simple \emph{heartbeat}-based technique where each parent node is
responsible for monitoring its children.

@@

\emph{Proximity in Kademlia} \cite{KLKP2008} introduces proximity
controls over the vanilla Kademlia protocol's random ID assignment, in order to
optimize the underlying network usage of the protocol. The work in this paper
can also be considered, as a more generic study over building proximity in the
context of iterative lookup algorithms. Focus is given on both the overlay and
the underlaying tiers in the sense that in the first improves the routing
performance according to a cost metric that is provided by the later and
quantifies the suitability of establishing a communication link between peers as
a cost function, based on the used proximity criteria (RTT, ISP locality, etc.).
%An underlay metric provides information about the underlay network. The paper
%  distinguishes between three kinds of possibilities to quantify it:
%\begin{itemize}
% \item Using measurements gained by previous lookups
% \item Using measurements acquired by other peers or jointly calculated with
%  others
% \item Using a local database to look up information
%\end{itemize}
The paper proposes two approaches for overlay optimization. One falling into
{\em proximity routing} and the other on {\em proximity neighbor selection}
approaches overlay optimization techniques for exploiting proximity. For the
first, the goal is in choosing the best next hop during routing a message. Due
to the fact that the routing in Kademlia is iterative, it is the initiator of
the lookup that chooses each next hop from a set of candidate nodes. Standard
Kademlia always chooses the closest node with respect to the XOR metric but with
proximity routing selection, Kademlia chooses the one with the smallest underlay
metric cost. Thus, as in all such approaches, there is a trade-off between the
overlay and underlay distances.

Towards the enhancement of traffic locality, the authors also use information
from the MaxMind GeoIP database\footnote{MaxMind Geolocation Technology.
http://www.maxmind.com} to retrieve proximity information of a given peer based
on its region, country, or ISP location, and use this information to form
clusters of nearby peers. An alternative method could also be the use of Vivaldi
coordinates \cite{cox_vivaldi_2004}. However, authors report that clustering
approach performs better than Vivaldi protocol based on their experiments.
Authors also report that {\em Proximity routing} protocol successfully worked
with the Kademlia protocol and improved the locality of the connections over
peers.

As already mentioned the paper proposes \emph{proximity neighbor selection}
method, as well, for overlay optimization. The goal, according to this, is to
keep peers with the least contact cost in the routing table. As Kademlia
constantly learns new peers (incoming requests, through iterative lookup) no
special algorithm for searching more cost effective peers is necessary and
simply choose the best peers seen so far.

@@

T2MC \cite{SLCGZ2008} uses traceroute logs to detect the AS boundaries and
cluster close by peers with each other to reduce the redundant multiple message
passes between the AS boundaries. T2MC uses a customized k-mean classification
algorithm with k = 2 to perform the classification, and exploits the stable
structure of the internet routers to guide clustering. Even though traceroute
provides detailed information about the network structure, use of traceroute
creates overhead to the overall network structure. For this same reason, it is
not unusual for network administrators to disable this support on their network
routers, which may affect the performance of the T2MC algorithm.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{PChord}
\emph{PChord}\cite{HLYW2005} is a based on the Chord DHT which adds
proximity routing into its routing mechanism. The main modification over the
vanilla Chord is the inclusion of a \emph{proximity list} into its routing table
so that the next hop is decided based not only by considering best progress
towards the key, but on the physical proximity of the candidate nodes as well.
The list is, at join time, empty but as the PChord node starts to interact with
other nodes it applies a heuristic mechanism to fill up the list. Entries are
dynamically added or removed as the network state changes. The overlay
maintenance cost can be considered similar to Chord's. Besides node join/leave
and cost of heartbeat messages for connection verification, PChord adds
heartbeat of nodes inside the proximity list but since this list always contains
nodes with the lowest RTT the added cost is kept to a minimal.

Due to the fact that, with each step, there is always progress towards the
target node in the ID space, PChord will result in a reduced hop number than
Chord, since each hop in PChord is larger or at least equal in the key space.
Passing through proximity links in the underlaying network means reduced routing
cost. Additionally, if the number of entries in the proximity list is the same
as the number of network partitions, PChord prevents hops from jumping in and
out the network partition the current node belongs to.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{AChord}
\emph{AChord} \cite{DK2006} uses IPv6's any-cast mechanism in order to
relieve the protocol from complex joining procedures, and achieve high accuracy
of network proximity giving high routing efficiency, Any-cast delivers a message
coming from the outside of an any-cast group to the physically closest node in
that group. AChord organizes all nodes participating in the overlay network into
an any-cast group. Any joining node comes outside of that group and, thus, is
automatically forwarded to the physically nearest node in order to bootstrap,
avoiding the need to, explicitly, maintain such nodes, and the effort of finding
a way of locating the physically nearest among them. The ID of the new coming
node is computed based on the ID of the bootstrap node and the bootstrap's
predecessor in a way that the constructed ID will be positioned between the
aforementioned two. After joining, the finger table is build the same way as in
the Chord protocol. Moreover, additional nodes are maintained in a structure
called \emph{adjacency table} which stores information about the closest known
node. The routing decision is made by using both the adjacency and the finger
table. The authors claim this to be a lightweight approach that requires few
changes to Chord in order not to affecting its own advantageous characteristics.

% TODO: MAYBE REVIEW THE EVALUATION
\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% good routing efficiency
medium &
% Relieves the system from joining overhead and topology aware ID construction
% but doubles the information needed to be stored and maintained by the
% introduction of adjacency table
medium &
% assumes the operation of ideal underlying any-cast mechanism (meaning that
% outside messages are always routed to the nearest node which in the current
% Internet, this is not always feasible. Moreover, is designed on the basis of
% IPv6 protocol which is not widely operational yet.
medium
\end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Chord6}
\emph{Chord6} \cite{XZHL2005} is another Chord variant that tries to
exploit the hierarchical features of IPv6 in order to create a substrate that
reduces inter-domain traffic between service providers. The difference between
Chord6 and Chord relies on the identifier definition. Therefore, the approach
can be easily portable to other DHTs such as CAN, Pastry and Tapestry. In Chord6
the identifier contains two parts: the higher bits are obtained by hashing the
node's IPv6 address prefix of specific length, while the remaining lower bits
are the hash value of the rest of that IPv6 address. As a result of this
assignment, nodes in a domain will be mapped onto a continuous key space on the
overlay network, which avoids unnecessary message forwarding across different
service providers, thus minimizing overall routing cost.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{DHT-PNS}
The work in \cite{DLTZZ2006} propose a proximity neighbor selection
scheme on top of the Chord DHT. The main goal is to use proximity information
to group physically close by nodes as neighbors in the DHT table. In order to
detect proximity, virtual network coordinates of peers are used, by using the
Vivaldi protocol \cite{cox_vivaldi_2004}. The virtual coordinates are then used
by the nodes to map to identifier space in the DHT. The space is partitioned
using a \emph{concentric circle clustering scheme} where successive cycles of
radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed. Then the formed
annuluses are divided into $2\chi-1$ \emph{sectors}, where $\chi$ denotes the
sequence number of the annulus starting from $\chi = 1$ for the center cycle. It
is proved in the paper, that this way each sector occupies the same area as does
the center cycle. Assuming uniform node distribution, this characteristic,
favors a more load balanced clustering operation. Every sector in this $2d$
coordinate space is mapped to a unique \emph{region} in the DHT space forming a
multi-layer node identifier space.  Thus, any nodes that belong to the same
sector, are mapped to the same region as well, preserving their proximity
relationship unveiled by the use of the Vivaldi protocol. The individual pieces
are mapped to the identifier space uniquely, allowing logarithmic lookup
operations with high probability on Chord. The system described in the previous
section allows any node to obtain its DHT key and its region key, in a fully
distributed manner, just by applying a consistent hash function. Using an RPC
call in Chord, the node can obtain the region's master node, called
\emph{Cluster Node (CN)} which is responsible for clustering the nodes belonging
in the same sector or region with that of the node at hand. An additional RPC
call, registers it to its corresponding region and publishes its information,
through the special peer CN. The node can, additionally, ask CN for other
nodes that have previously joined the region in order to add them into its
neighbor set for future routing table optimization. Even in the case when no
neighbor is detected in the current region, the search is expanded to adjacent
regions and towards an upper layer identifier space until one is found or the
first layer reached.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{IP-based Clustering (IPBC)}
Proximity neighbor selection algorithms use probing and other techniques to
detect proximity. However, such methods are either not precise enough, or
overload the network. \emph{IP-based clustering} \cite{KM2007} is a proximity
neighbor selection based algorithm in which the authors propose to use the IP
address prefixes (16 bit for IPv4) in order to detect proximity.
\cite{freedman_iploc_2005} states that $97\%$ of prefixes larger than $24$ bits
belong to a single geographical location. However, using less number of bits
creates less precise results and more number of bits increase the burden on the
network and reduce the possible number of neighbors. Therefore, a careful
selection is required in terms of performance/accuracy trade-off. IP-based
Clustering generates a key and its prefix and stores the later in the DHT
itself, so that any newly joining nodes with the same IP prefix can query it and
identify all the neighbors with the same prefix relatively easy. Nodes
periodically update their entries in the DHT by removing those who leave, or
those who has been timed out with no further activity detected from the node.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Cone}
\emph{Cone} \cite{HY2007} extends Chord using proximity neighbor
selection topology optimization algorithm. The proximity information is
generated using landmarks and round-trip-time-based distances to landmarks.
Cone uses a two-layered identifier space. The first, named Chord-layer
identifier, denoted as $Id_{Chord}$, is the same as in Chord. The second is the
Cone-layer identifier, $Id_{Cone}$ which is constructed by two component
identifiers. The first, known as \emph{group identifier (gid)} denotes a
relevant group the node belongs to while the second, namely \emph{local
identifier (lid)} indicates the local identifier within the group. The group
concept, which is introduced here, is a way of dividing nodes according to a
common $Id_{Chord}$ prefix.  The structure of a Cone overlay, retains the
Chord's circular topology. The difference lies on the fact that, now, two rings
are created. A big ring, where nodes with the same \emph{gid} are arranged at
each position. Each of these positions are a smaller ring for the particular
group's \emph{lid}s. The routing is achieved in both clockwise and
counter-clockwise directions in the big-ring, for which two routing tables are
maintained, namely \emph{front} and \emph{back} finger tables. Entries in these
tables, display physical network proximity with the current node. Moreover, a
third table called \emph{group} maintains information about other on-line peers
within the current node's group in a way that entries are now close in the ID
space.

@@

A quick overview can be obtained in Table~\ref{structured:table}. Topology
adaptation, landmarking and caching/replication have been visited during the
discussion on unstructured systems in Section~\ref{section:unstructured} so
we will not elaborate more here. Again, we additionally gather highlights and a
rough estimation of each algorithm's positive and negative aspects regarding
their implementation.

@@

\begin{figure}[ht]
\centering
  \includegraphics[scale=0.4]{img/pdf/proximity-routing.pdf}
\caption{Proximity routing takes proximity info into account during the
routing process.}
\label{figure:proximity-routing}
\end{figure}

>>>>

\section{Structured P2P Networks}
\label{section:structured}

This section presents a description and analysis of algorithms proposed
to address the topology mismatch problem in structured P2P networks.
In structured P2P networks, the construction of the routing tables
held by participating nodes determines the efficiency of message forwarding.
Routing tables that closely match the underlying IP topology
achieve increased performance. As a result, the proposed solutions for
the topology mismatch problem for structured P2P networks 
use different levels of proximity
information to optimize the routing tables used by peers.  
The categorization of the algorithms we present here is based on the
categorization suggested by previous 
work~\cite{CDHR2002,CDCR2002,RSS2002}.

@@

\textit{Global Soft-State} \cite{XTZ2003} builds a global map to help
choose shorter routing paths, combining the landmark binning method and small
scale distance probes to reveal the proximity properties of the underlying
network to the overlay. %Figure \ref{fig:global_softstate} presents the landmark
%space and the mapping of the position of the nodes.
This global view of the
state is made available to all nodes to help them find the best way to
route their messages. \textit{Global Soft-State} operates in two phases: 
generation and use of the proximity information. For generating the proximity
information a hybrid approach is proposed, which uses landmark clustering as a
preprocessing step to select a number of potential nearest neighbor
candidates and then refine the selection by incorporating a
round-trip-time-based scheme to ultimately choose the closest node. For using
the proximity information, the algorithm chooses a different path from the
classic gossiping approaches for constructing and maintaining the overlay. It is
based on landmark clustering for strategic placement of proximity information on
the overlay enabling any node to access such information using a landmark number
that reflects its physical position in the network. For various logical
regions\footnote{This might be a high-order zone in the eCAN\cite{xu_ecan_2002}
context or a set of nodes sharing a particular prefix in overlays such as
Pastry\cite{antony_pastry_2001}.} maps of physical information are built and
published where each node may appear in a maximum of $log\left( N \right)$ such
maps. To dynamically adapt to changing network conditions, a node subscribes to
relevant \emph{soft states} that utilize a notification system in order to
initialize any necessary neighbor re-selection.

Maintaining several host states at different layers, makes any content migration
costly. Additionally, the method does not make any continuing effort to remap
the overlay structure after a node successfully joins in order to adapt its
state to any condition change. Although this approach greatly reduces the
routing latency to far nodes, it is unable to dynamically identify nodes that
are close to routers and gateways in order to construct the secondary overlay.
Nevertheless, static recognition of such nodes is currently done based on BGP
reports and pre-chosen landmarks, sacrificing the self-organizing attribute of
traditional DHTs.

@@

The bootstrap phase of Mithos starts with a subset of existing members as the
first set of candidates, and while the iteratively closest nodes are detected by
probing the neighborhood, the candidate neighbor list is updated. To
avoid a local minima, Mithos probes all the neighbors within a two-hop distance
from the current minimum before concluding the process.

After finding its first neighbor, the newcoming peer is assigned an ID using
information gathered during the iterative neighbor selection phase. Virtual
coordinates are assigned to the newcoming peer by using the distances of the two
closest nodes and their neighbors, so that Euclidean distances between the node
and all known hosts predict the network latency between them
\cite{cox_vivaldi_2004}. The benefit of these synthetic coordinates are that
they are explicitly used as the node's ID, and distance computations among nodes
can be done in ID space without requiring physical measurements.

The last step of the algorithm is the interconnection among neighbors. Mithos
uses a \emph{quadrant-based} mechanism in which each node establishes
a link to the closest neighbor in each quadrant. During forwarding, the next
hop is performed towards a neighbor in the same quadrant as the final
destination. The new coming node may not know of other neighbors in all
quadrants, therefore, the node first identifies neighbors in all quadrants
using a mechanism based on ideas similar to a perimeter walk\footnote{Used in
the Greedy Perimeter Stateless Routing (GPSR) protocol.} and then improves the
results using parallel path processing by taking into account further geometric
properties of node relationships.

One major limitation of \textit{Mithos} is that the protocol cannot handle
dynamic arrivals and departures from the network.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{LAPTOP}
\emph{LAPTOP} \cite{WLH2007}, organizes the overlay into a tree-based hierarchy
with the goal of reducing hops during message routing as well as minimizing
maintenance overhead. Additionally, a caching scheme is also incorporated so as
to further reduce routing table update costs. The authors theoretically show
that in LAPTOP, routing path length is bounded by $O(log_d N)$ and node joining
and leaving in the overlay network is bounded by $O\left( d log_d N \right)$
hops in a balanced overlay tree, where $N$ is the number of nodes, and $d$ is
the maximum degree of each node. LAPTOP implements the geographical layout
approach  and constructs its layout in a self-organizing and efficient fashion.
The physical distances are estimated with round trip times to a few existing
nodes in the overlay network.

Each node is assigned a dot-formated address (e.g 1.3.4). Each octet ranges from
$1$ to $d$, where $d$ is the maximum degree of the nodes. The assignment process
is done by appending a unique octet to the address of each node's parent, while
the root node is assigned address $1$.

@@

The routing scheme is similar to the longest-prefix IP matching scheme. At each
forwarding hop, any message travels up the tree until the first common ancestor
of both source and destination nodes is reached and then starts descending to
arrive to its target. During tree traversals, special entries in the routing
tables, called \emph{routing cache}, are maintained to increase routing
efficiency and achieve finer load balance. Caching enables a node to forward a
message to a better longest-prefix match than that of its direct ancestor making
a larger, quicker and more cost effective step through the overlay and toward
the destination. To improve scalability, the number of child nodes and the
size of the routing cache are limited. In terms of overlay maintenance, LAPTOP
incorporates a simple \emph{heartbeat}-based technique where each parent node is
responsible for monitoring its children.

@@

\emph{Proximity in Kademlia} \cite{KLKP2008} introduces proximity
controls over the vanilla Kademlia protocol's random ID assignment to
optimize the underlying network usage of the protocol. The work in this paper
can also be considered as a more generic study of incorporating proximity into
iterative lookup algorithms. 
The overlay tier improves its routing
performance using a cost metric that is provided by the underlying tier
and quantifies the suitability of establishing a communication link between peers as
a cost function, based on the used proximity criteria (RTT, ISP locality, etc.).
%An underlay metric provides information about the underlay network. The paper
%  distinguishes between three kinds of possibilities to quantify it:
%\begin{itemize}
% \item Using measurements gained by previous lookups
% \item Using measurements acquired by other peers or jointly calculated with
%  others
% \item Using a local database to look up information
%\end{itemize}

The paper proposes two approaches for overlay optimization: 
{\em proximity routing} and {\em proximity neighbor selection}.
In proximity routing, the goal
is to choose the best next hop during routing a message. Due
to the fact that the routing in Kademlia is iterative, it is the initiator of
the lookup that chooses each next hop from a set of candidate nodes. Standard
Kademlia always chooses the closest node with respect to the XOR metric but with
proximity routing selection, Kademlia chooses the node with the smallest underlay
metric cost. Thus, as in all such approaches, there is a trade-off between the
overlay and underlay distances.

Towards the enhancement of traffic locality, the authors also use information
from the MaxMind GeoIP database\footnote{MaxMind Geolocation Technology.
http://www.maxmind.com} to retrieve proximity information of a given peer based
on its region, country, or ISP location, and use this information to form
clusters of nearby peers. An alternative method could use Vivaldi
coordinates \cite{cox_vivaldi_2004}. However, the authors report that their clustering
approach performs better than the Vivaldi protocol based on their experiments.
Authors also report that {\em Proximity routing} protocol successfully worked
with the Kademlia protocol and improved the locality of the connections between
peers.

In \emph{proximity neighbor selection}, the goal is to
keep peers with the least contact cost in the routing table. As Kademlia
continuously learns about new peers via incoming requests and an iterative
lookup process, no 
special algorithm for searching more cost-effective peers is necessary and
it simply chooses the best peers seen so far.

@@

T2MC \cite{SLCGZ2008} uses traceroute logs to detect the AS boundaries and
cluster nearby peers together to reduce the redundant multiple message
passes between the AS boundaries. T2MC uses a customized k-mean classification
algorithm with k = 2 to perform the classification, and exploits the stable
structure of the internet routers to guide clustering. Even though traceroute
provides detailed information about the network structure, use of traceroute
creates overhead on the overall network structure. For this same reason, it is
not unusual for network administrators to disable this support on their network
routers, which may affect the performance of the T2MC algorithm.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{PChord}
\emph{PChord}\cite{HLYW2005} is based on the Chord DHT and adds
proximity routing into its routing mechanism. The main modification over the
vanilla Chord is the inclusion of a \emph{proximity list} into its routing table
so that the next hop is decided based on both the best progress
towards the key and the physical proximity of the candidate nodes as well.
The list is, at join time, empty but as the PChord node starts to interact with
other nodes it applies a heuristic mechanism to fill the list. Entries are
dynamically added or removed as the network state changes. The overlay
maintenance cost can be considered similar to Chord's. Besides node join/leave
and cost of heartbeat messages for connection verification, PChord adds
heartbeat of nodes inside the proximity list but since this list always contains
nodes with the lowest RTT the added cost is kept to a minimum.

Due to the fact that, with each step, there is always progress towards the
target node in the ID space, PChord will result in a reduced hop number than
Chord, since each hop in PChord is larger or at least equal in the key space.
Passing through proximity links in the underlying network means reduced routing
cost. Additionally, if the number of entries in the proximity list is the same
as the number of network partitions, PChord prevents hops from jumping in and
out of the network partition the current node belongs to.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{AChord}
\emph{AChord} \cite{DK2006} uses IPv6's anycast mechanism to
relieve the protocol from complex joining procedures, and achieve high accuracy
of network proximity and thus, high routing efficiency. Anycast delivers a message
coming from the outside of an anycast group to the physically closest node in
that group. AChord organizes all nodes participating in the overlay network into
an anycast group. Any joining node comes outside of that group and, thus, is
automatically forwarded to the physically nearest node in order to bootstrap,
avoiding the need to, explicitly, maintain such nodes, and the effort of finding
a way of locating the physically nearest among them. The ID of the newcoming
node is computed based on the ID of the bootstrap node and the bootstrap's
predecessor in a way that the constructed ID will be positioned between the
aforementioned two. After joining, the finger table is built the same way as in
the Chord protocol. Moreover, additional nodes are maintained in a structure
called \emph{adjacency table} which stores information about the closest known
node. The routing decision is made by using both the adjacency and the finger
table. The authors claim this to be a lightweight approach that requires few
changes to Chord while keeping its advantageous characteristics.

% TODO: MAYBE REVIEW THE EVALUATION
\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% good routing efficiency
medium &
% Relieves the system from joining overhead and topology aware ID construction
% but doubles the information needed to be stored and maintained by the
% introduction of adjacency table
medium &
% assumes the operation of ideal underlying anycast mechanism (meaning that
% outside messages are always routed to the nearest node which in the current
% Internet, this is not always feasible. Moreover, is designed on the basis of
% IPv6 protocol which is not widely operational yet.
medium
\end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Chord6}
\emph{Chord6} \cite{XZHL2005} is another Chord variant that tries to
exploit the hierarchical features of IPv6 to create a substrate that
reduces inter-domain traffic between service providers. The difference between
Chord6 and Chord relies on the identifier definition. Therefore, the approach
can be easily portable to other DHTs such as CAN, Pastry and Tapestry. In Chord6
the identifier contains two parts: the higher bits are obtained by hashing the
node's IPv6 address prefix of specific length, while the remaining lower bits
are the hash value of the rest of the IPv6 address. As a result of this
assignment, nodes in a domain will be mapped onto a continuous key space on the
overlay network, which avoids unnecessary message forwarding across different
service providers, thus minimizing overall routing cost.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{DHT-PNS}
The work in \cite{DLTZZ2006} propose a proximity neighbor selection
scheme on top of the Chord DHT. The main goal is to use proximity information
to group physically close by nodes as neighbors in the DHT table. In order to
detect proximity, virtual network coordinates of peers are used, by using the
Vivaldi protocol \cite{cox_vivaldi_2004}. The virtual coordinates are then used
by the nodes to map to identifier space in the DHT. The space is partitioned
using a \emph{concentric circle clustering scheme} where successive cycles of
radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed. Then the formed
annuluses are divided into $2\chi-1$ \emph{sectors}, where $\chi$ denotes the
sequence number of the annulus starting from $\chi = 1$ for the center cycle. It
is proved in the paper, that this way each sector occupies the same area as does
the center cycle. Assuming uniform node distribution, this characteristic,
favors a more load balanced clustering operation. Every sector in this $2d$
coordinate space is mapped to a unique \emph{region} in the DHT space forming a
multi-layer node identifier space.  Thus, any nodes that belong to the same
sector, are mapped to the same region as well, preserving their proximity
relationship unveiled by the use of the Vivaldi protocol. The individual pieces
are mapped to the identifier space uniquely, allowing logarithmic lookup
operations with high probability on Chord. The system described in the previous
section allows any node to obtain its DHT key and its region key, in a fully
distributed manner, just by applying a consistent hash function. Using an RPC
call in Chord, the node can obtain the region's master node, called
\emph{Cluster Node (CN)} which is responsible for clustering the nodes belonging
in the same sector or region with that of the node at hand. An additional RPC
call, registers it to its corresponding region and publishes its information,
through the special peer CN. The node can, additionally, ask CN for other
nodes that have previously joined the region to add them into its
neighbor set for future routing table optimization. Even in the case when no
neighbor is detected in the current region, the search is expanded to adjacent
regions and towards an upper layer identifier space until one is found or the
first layer reached.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{IP-based Clustering (IPBC)}
Proximity neighbor selection algorithms use probing and other techniques to
detect proximity. However, such methods are either not precise enough, or
overload the network. \emph{IP-based clustering} \cite{KM2007} is a proximity
neighbor selection based algorithm in which the authors propose to use the IP
address prefixes (16 bit for IPv4) in order to detect proximity.
\cite{freedman_iploc_2005} states that $97\%$ of prefixes larger than $24$ bits
belong to a single geographical location. However, using a smaller number of bits
creates less precise results and a larger number of bits increases the burden on the
network and reduces the possible number of neighbors. Therefore, a careful
selection is required in terms of performance/accuracy trade-off. IP-based
clustering generates a key and its prefix and stores the prefix in the DHT
itself, so that any newly joining nodes with the same IP prefix can query it and
identify all the neighbors with the same prefix easily.  Nodes
periodically update their entries in the DHT by removing those who leave, or
those who have been timed out with no further activity detected from the node.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Cone}
\emph{Cone} \cite{HY2007} extends Chord using a proximity neighbor
selection topology optimization algorithm. The proximity information is
generated using landmarks and round-trip-time-based distances to landmarks.
Cone uses a two-layered identifier space. The first, named Chord-layer
identifier, denoted as $Id_{Chord}$, is the same as in Chord. The second is the
Cone-layer identifier, $Id_{Cone}$ which is constructed using two component
identifiers. The first, known as \emph{group identifier (gid)} denotes a
relevant group to which the node belongs while the second, namely \emph{local
identifier (lid)} indicates the local identifier within the group. The group
concept is a way of dividing nodes according to a
common $Id_{Chord}$ prefix.  The structure of a Cone overlay, retains the
Chord's circular topology. The difference lies in the fact that, now, two rings
are created. A big ring, where nodes with the same \emph{gid} are arranged at
each position. Each of these positions are a smaller ring for the particular
group's \emph{lid}s. Routing is achieved in both clockwise and
counter-clockwise directions in the big-ring, for which two routing tables are
maintained, namely \emph{front} and \emph{back} finger tables. Entries in these
tables, display physical network proximity with the current node. Moreover, a
third table called \emph{group} maintains information about other on-line peers
within the current node's group in a way that entries are now close in the ID
space.

@@

A quick overview can be obtained in Table~\ref{structured:table}. Topology
adaptation, landmarking and caching/replication have been visited during the
discussion on unstructured systems in Section~\ref{section:unstructured} so
we will not elaborate more here. As in the previous section, we provide the 
key properties of each algorithm that differentiate it from the others and
highlight 
each algorithm's advantages and disadvantages.

@@

\begin{figure}[ht]
\centering
  \includegraphics[scale=0.4]{img/pdf/proximity-routing.pdf}
\caption{Proximity routing takes proximity information into account during the
routing process.}
\label{figure:proximity-routing}
\end{figure}
